### 2020-03-06 Common/chapters/introduction.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

\chapter{Introduction}
\label{ch:introduction}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}. An early detection and treatment increases the chances of recovery. In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses. Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.

@@

\section{Motivation}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Cancer is a serious concern in today's society. An early detection is key to maximize the chances of recovery. As explained by Nicholas Petrick~\cite{50}, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device. The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information". Under these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians. A complete CAD system is composed of two parts. The first one is called CADe for "computer-aided detection". This part includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection. The other part is CADx for "computer-aided diagnosis" which aims at extracting the characteristics of lesions to classify them according to their malignancy. CAD systems provide multiple advantages.\\
First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses. Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.\\
Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection. As stated before, detecting cancers at an early stage maximizes the probability of curing them. Another benefit of using CAD systems, especially the ones based on deep learning, is the number of cases that the latter are trained on. In fact, in order to built efficient deep learning models, large amounts of data are required. Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.\\
However, these advantages remain partially theoretical nowadays. Research is still ongoing, aiming at improving existent CAD systems. Problems like poor detection rates, small amounts of available data and bad levels of generalization are still recurrent. Therefore, this situation can and must be improved.

@@

\section{Contributions}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field. Concretely, the major contributions of this work are:
\begin{itemize}
\item A state-of-the-art deep learning model for prostate cancer detection. All steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented. This makes the experiment completely reproductible.

>>>>

\chapter{Introduction}
\label{ch:introduction}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}
According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}. An early detection and treatment increases the chances of recovery. In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses. Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.

@@

\section{Motivation}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Cancer is a serious concern in today's society. An early detection is key to maximize the chances of recovery. As explained by Nicholas Petrick~\cite{50}, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device. The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information". Under these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians. A complete CAD system is composed of two parts. The first one is called CADe for "computer-aided detection". This part includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection. The other part is CADx for "computer-aided diagnosis" which aims at extracting the characteristics of lesions to classify them according to their malignancy. CAD systems provide multiple advantages.

First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses. Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.

Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection. As stated before, detecting cancers at an early stage maximizes the probability of curing them. Another benefit of using CAD systems, especially the ones based on deep learning, is the number of cases that the latter are trained on. In fact, in order to built efficient deep learning models, large amounts of data are required. Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.

However, these advantages remain partially theoretical nowadays. Research is still ongoing, aiming at improving existent CAD systems. Problems like poor detection rates, small amounts of available data and bad levels of generalization are still recurrent. Therefore, this situation can and must be improved.

@@

\section{Contributions}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field. Concretely, the major contributions of this work are:
\begin{itemize}
\item A state-of-the-art deep learning model for prostate cancer detection. All steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented. This makes the experiment completely reproductible.

### 2020-03-06 Common/chapters/literature_review.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

\chapter{Literature review}
\label{ch:literature_review}
%Traditionally, building a CAD system consisted of two different steps: "Lesion  detection (CADe) and  false-positive  reduction (CADx).  Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions. False-positive reduction was commonly based on traditional  machine  learning  methods  to  reduce  the  false  positive  rate. However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"~\cite{41}.\\
%Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks. In fact, the latter can be considered as single-step CAD system. For this reason, CAD systems are currently a trending topic in deep learning.\\

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e. prostate, lung and brain. Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally. It was predicted that there would be 1.7 million new cancer cases by 2030. The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}. Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection. All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see Section \ref{sec:prostatex_dataset_description}).

Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs. Their data processing approach kept T2W, DWI and ADC grayscale images only. After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65$x$65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels. Thanks to this method, the same lesion is visible in the same area over the three channels. This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter. Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e. by subtracting the mean before dividing by the standard deviation. The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$). The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1. Most of the processing techniques are reproductible, apart from the manual lesion contouring and labelling performed by a radiologist. 
Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}). The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set. The best results were obtained by using DWI images only with the highest b-value, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$). However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.

@@

\section{Lung/Lung CT Challenge}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Gao et al. claimed that "lung cancer is one of the most frequent and leading causes of death all over the world. It was reported that there were approximately $1.8$x$10^6$ new cases of lung cancer globally in 2012. Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}. The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).

@@

Armato et al.~\cite{12} described the LUNGx Challenge and the overall results. This challenge consisted in classifying the lung nodules as benign or malignant among a training set of 10 scans and a test set of 60 scans. Since the training set was extremely small, training the model on other datasets was allowed. The article describes the results of the proposed methods and compares them with the performance of six qualified radiologists on the same task. The article reports that "ten groups applied their own methods to 73 lung nodules (37 benign and 36 malignant) that were selected to achieve approximate size matching between the two cohorts. Area under the receiver operating characteristic curve (AUC) values for these methods ranged from 0.50 to 0.68. Only three methods performed statistically better than random guessing. The radiologists’ AUC values ranged from 0.70 to 0.85. Three radiologists performed statistically better than the best-performing computer method."~\cite{12}. Figure \ref{fig:LUNGx_challenge_all_results} gives an overview of the results.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Brain cancer is another major type of cancer. According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women. It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}. The following research papers are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website \cite{45}. However, since it is the only dataset available for brain tumor classification, some publications used it. Further details about this dataset can be found in section \ref{sec:kaggleBrain}.

@@

%Hanwat et al.~\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset. Their processing method included skull masking, i.e. the removal of every non-brain tissue from the image. According to the authors, this technique improves performance quite a lot. ---> Check their methods (graphs look special)

Habib Mohamed Ali~\cite{04} proposed his own convolutional neural network to classify the images of the Kaggle Brain dataset. First, the data was augmented. Then, it was cropped thanks to OpenCV so that the resulting images only contained the brain itself. Afterwards, they were resized and normalized in order to scale pixel values to the range $[0,1]$. Once the processing part was over, the data was split into a training set ($70\%$), a validation set ($15\%$) and a test set ($15\%$). Regarding the neural network structure, the model is simple. It is composed of only one convolutional layer with a batch normalization layer and ReLU activation function, followed by two max-pooling layers and a dense layer. This model achieved an accuracy of $89\%$ on the test set.

>>>>

\chapter{Literature review}
\label{ch:literature_review}
%Traditionally, building a CAD system consisted of two different steps: "Lesion  detection (CADe) and  false-positive  reduction (CADx).  Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions. False-positive reduction was commonly based on traditional  machine  learning  methods  to  reduce  the  false  positive  rate. However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"~\cite{41}.\\
%Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks. In fact, the latter can be considered as single-step CAD system. For this reason, CAD systems are currently a trending topic in deep learning.\\
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e. prostate, lung and brain. Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally. It was predicted that there would be 1.7 million new cancer cases by 2030. The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}. Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection. All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is composed of prostate lesions multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans)(see Section \ref{sec:prostatex_dataset_description}).

Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs. Their data processing approach kept T2W, DWI and ADC grayscale images only. After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65$x$65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels. Thanks to this method, the same lesion is visible in the same area over the three channels. This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter. Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e. by subtracting the mean before dividing by the standard deviation. The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$). The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}). The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set. The best results were obtained by using DWI images only with the highest b-value, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$). However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.

@@

\section{Lung/Lung CT Challenge}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Gao et al. claimed that "lung cancer is one of the most frequent and leading causes of death all over the world. It was reported that there were approximately $1.8$x$10^6$ new cases of lung cancer globally in 2012. Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}. The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).

@@

Armato et al.~\cite{12} described the LUNGx Challenge and the overall results. This challenge consisted in classifying the lung nodules as benign or malignant among a training set of 10 scans and a test set of 60 scans. Since the training set was extremely small, training the model on other datasets was allowed. The article describes the results of the proposed methods and compares them with the performance of six qualified radiologists on the same task. The article reports that "ten groups applied their own methods to 73 lung nodules (37 benign and 36 malignant) that were selected to achieve approximate size matching between the two cohorts. Area under the receiver operating characteristic curve (AUC) values for these methods ranged from 0.50 to 0.68. Only three methods performed statistically better than random guessing. The radiologists’ AUC values ranged from 0.70 to 0.85. Three radiologists performed statistically better than the best-performing computer method."~\cite{12}. Figure \ref{fig:LUNGx_challenge_all_results} reports all methods used to classify the lung lesions with their associated performance. Furthermore, it also provides the performance of all radiologists.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Brain cancer is another major type of cancer. According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women. It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}. The following research papers are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website \cite{45}. However, as it is the only dataset available for brain tumor classification, some publications used it. Further details can be found in section \ref{sec:kaggleBrain}.

@@

%Hanwat et al.~\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset. Their processing method included skull masking, i.e. the removal of every non-brain tissue from the image. According to the authors, this technique improves performance quite a lot. ---> Check their methods (graphs look special)
Habib Mohamed Ali~\cite{04} proposed his own convolutional neural network to classify the images of the Kaggle Brain dataset. First, the data was augmented. Then, it was cropped so that the resulting images only contained the brain itself. Afterwards, they were resized and normalized in order to scale pixel values to the range $[0,1]$. Once the processing part was over, the data was split into a training set ($70\%$), a validation set ($15\%$) and a test set ($15\%$). Regarding the neural network structure, the model is simple. It is composed of only one convolutional layer with a batch normalization layer and ReLU activation function, followed by two max-pooling layers and a dense layer. This model achieved an accuracy of $89\%$ on the test set.
